Palaeobiology demo: disparity-through-time and within groups

### 2018-03-01 inst/gitbook/02_getting-started.Rmd
Changed time.subsets and merge.subsets to chrono.subsets and combine.subsets

Note that `disparity.through.time` and `disparity.per.group` are wrapper functions (i.e. they incorporate lots of other functions) that allow users to run a basic disparity-through-time, or disparity among groups, analysis without too much effort.
As such they use a lot of default options. 
These are described in the help files for the functions that are used to make the wrapper functions, and not described in the help files for `disparity.through.time` and `disparity.per.group`. 
These defaults are good enough for **data exploration**, but for a proper analysis you should consider the **best parameters for your question and data**. 
For example, which metric should you use?
How many bootstraps do you require?
What model of evolution is most appropriate if you are time slicing?
Should you rarefy the data?
See [`time.subsets`](#time-slicing), [`custom.subsets`](#customised-subsets), [`boot.matrix`](#bootstraps-and-rarefactions) and [`dispRity.metric`](#disparity-metrics) for more details of the defaults used in each of these functions.
Note that any of these default arguments can be changed within the `disparity.through.time` or `disparity.per.group` functions.

>>>>

Note that `disparity.through.time` and `disparity.per.group` are wrapper functions (i.e. they incorporate lots of other functions) that allow users to run a basic disparity-through-time, or disparity among groups, analysis without too much effort.
As such they use a lot of default options. 
These are described in the help files for the functions that are used to make the wrapper functions, and not described in the help files for `disparity.through.time` and `disparity.per.group`. 
These defaults are good enough for **data exploration**, but for a proper analysis you should consider the **best parameters for your question and data**. 
For example, which metric should you use?
How many bootstraps do you require?
What model of evolution is most appropriate if you are time slicing?
Should you rarefy the data?
See [`chrono.subsets`](#time-slicing), [`custom.subsets`](#customised-subsets), [`boot.matrix`](#bootstraps-and-rarefactions) and [`dispRity.metric`](#disparity-metrics) for more details of the defaults used in each of these functions.
Note that any of these default arguments can be changed within the `disparity.through.time` or `disparity.per.group` functions.

### 2018-03-01 inst/gitbook/03_specific-tutorials.Rmd
Changed time.subsets and merge.subsets to chrono.subsets and combine.subsets

The function `time.subsets` allows users to divide the matrix into different time subsets or slices given a dated phylogeny that contains all the elements (i.e. taxa) from the matrix.
Each subset generated by this function will then contain all the elements present at a specific point in time or during a specific period in time.

@@

```{r, eval=TRUE}
## Generating three time bins containing the taxa present every 40 Ma
time.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0))
```

@@

```{r, eval=TRUE}
## Automatically generate three equal length bins:
time.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = 3)
```

@@

## Generating time bins including taxa that might span between them
time.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0), FADLAD = BeckLee_ages)
```

@@

```{r, eval=TRUE}
## Generating four time slices every 40 million years under a model of proximity evolution
time.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)

## Generating four time slices automatically
time.subsets(data = BeckLee_mat99, tree = BeckLee_tree,
    method = "continuous", model = "proximity", time = 4, FADLAD = BeckLee_ages)
```

@@

Of course, one could directly supply the subsets generated above (using `time.subsets` or `custom.subsets`) to this function.

@@

## Creating time slice subsets
time_slices <- time.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
                               method = "continuous", model = "proximity", 
                               time = c(120, 80, 40, 0),
                               FADLAD = BeckLee_ages)

@@

## Creating time slice subsets
time_slices <- time.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)
## Bootstrapping the time slice subsets
boot_time_slices <- boot.matrix(time_slices, bootstraps = 100)
## Calculate disparity
disparity_time_slices <- dispRity(boot_time_slices, metric = c(sum, variances))

## Creating time bin subsets
time_bins <- time.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "discrete", time = c(120, 80, 40, 0), FADLAD = BeckLee_ages,
    inc.nodes = TRUE)
## Bootstrapping the time bin subsets
boot_time_bins <- boot.matrix(time_bins, bootstraps = 100)
## Calculate disparity
disparity_time_bins <- dispRity(boot_time_bins, metric = c(sum, variances))
```

@@

## Plotting the results with some plot.dispRity arguments
plot(disparity_time_slices, quantile = c(seq(from = 10, to = 100, by = 10)),
    cent.tend = sd, type = "c", elements = TRUE, col = c("black", rainbow(10)),
    ylab = c("Disparity", "Diversity"), time.subsets = FALSE,
    xlab = "Time (in in units from past to present)", observed = TRUE,
    main = "Many more options...")

@@

Finally, it is possible to use objects generated by `time.subsets`.
In this case, `adonis.dispRity` will applied the `matrix ~ time` formula by default:

```{r}
## Creating time series
time_subsets <- time.subsets(BeckLee_mat50, BeckLee_tree, 
     method = "discrete", inc.nodes = FALSE, time = c(100, 85, 65, 0),
     FADLAD = BeckLee_ages)

@@

To use each time subset as a separate predictor, you can use the `matrix ~ time.subsets` formula; this is equivalent to `matrix ~ first_time_subset + second_time_subset + ...`:

```{r}
## Running the NPMANOVA with each time bin as a predictor
adonis.dispRity(time_subsets, matrix ~ time.subsets)
```

>>>>

The function `chrono.subsets` allows users to divide the matrix into different time subsets or slices given a dated phylogeny that contains all the elements (i.e. taxa) from the matrix.
Each subset generated by this function will then contain all the elements present at a specific point in time or during a specific period in time.

@@

```{r, eval=TRUE}
## Generating three time bins containing the taxa present every 40 Ma
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0))
```

@@

```{r, eval=TRUE}
## Automatically generate three equal length bins:
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = 3)
```

@@

## Generating time bins including taxa that might span between them
chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree, method = "discrete",
                time = c(120, 80, 40, 0), FADLAD = BeckLee_ages)
```

@@

```{r, eval=TRUE}
## Generating four time slices every 40 million years under a model of proximity evolution
chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)

## Generating four time slices automatically
chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree,
    method = "continuous", model = "proximity", time = 4, FADLAD = BeckLee_ages)
```

@@

Of course, one could directly supply the subsets generated above (using `chrono.subsets` or `custom.subsets`) to this function.

@@

## Creating time slice subsets
time_slices <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
                               method = "continuous", model = "proximity", 
                               time = c(120, 80, 40, 0),
                               FADLAD = BeckLee_ages)

@@

## Creating time slice subsets
time_slices <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "continuous", model = "proximity", time = c(120, 80, 40, 0),
    FADLAD = BeckLee_ages)
## Bootstrapping the time slice subsets
boot_time_slices <- boot.matrix(time_slices, bootstraps = 100)
## Calculate disparity
disparity_time_slices <- dispRity(boot_time_slices, metric = c(sum, variances))

## Creating time bin subsets
time_bins <- chrono.subsets(data = BeckLee_mat99, tree = BeckLee_tree, 
    method = "discrete", time = c(120, 80, 40, 0), FADLAD = BeckLee_ages,
    inc.nodes = TRUE)
## Bootstrapping the time bin subsets
boot_time_bins <- boot.matrix(time_bins, bootstraps = 100)
## Calculate disparity
disparity_time_bins <- dispRity(boot_time_bins, metric = c(sum, variances))
```

@@

## Plotting the results with some plot.dispRity arguments
plot(disparity_time_slices, quantile = c(seq(from = 10, to = 100, by = 10)),
    cent.tend = sd, type = "c", elements = TRUE, col = c("black", rainbow(10)),
    ylab = c("Disparity", "Diversity"), chrono.subsets = FALSE,
    xlab = "Time (in in units from past to present)", observed = TRUE,
    main = "Many more options...")

@@

Finally, it is possible to use objects generated by `chrono.subsets`.
In this case, `adonis.dispRity` will applied the `matrix ~ time` formula by default:

```{r}
## Creating time series
time_subsets <- chrono.subsets(BeckLee_mat50, BeckLee_tree, 
     method = "discrete", inc.nodes = FALSE, time = c(100, 85, 65, 0),
     FADLAD = BeckLee_ages)

@@

To use each time subset as a separate predictor, you can use the `matrix ~ chrono.subsets` formula; this is equivalent to `matrix ~ first_time_subset + second_time_subset + ...`:

```{r}
## Running the NPMANOVA with each time bin as a predictor
adonis.dispRity(time_subsets, matrix ~ chrono.subsets)
```

### 2018-03-01 inst/gitbook/06_palaeo-demo.Rmd
Changed time.subsets and merge.subsets to chrono.subsets and combine.subsets

The `dispRity` package provides a `time.subsets` function that allows users to split the morphospace into time slices (using `method = continuous`) or into time bins (using `method = discrete`).
In this example, we are going to split the morphospace into five equal time bins of 20 million years long from 100 million years ago to the present.
We will also provide to the function a table containing the first and last occurrences dates for some fossils to take into account that some fossils might occur in several of our different time bins.

@@

## Splitting the morphospace using the time.subsets function
(binned_morphospace <- time.subsets(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages))
```

>>>>

The `dispRity` package provides a `chrono.subsets` function that allows users to split the morphospace into time slices (using `method = continuous`) or into time bins (using `method = discrete`).
In this example, we are going to split the morphospace into five equal time bins of 20 million years long from 100 million years ago to the present.
We will also provide to the function a table containing the first and last occurrences dates for some fossils to take into account that some fossils might occur in several of our different time bins.

@@

## Splitting the morphospace using the chrono.subsets function
(binned_morphospace <- chrono.subsets(data = BeckLee_mat50, tree = BeckLee_tree,
    method = "discrete", time = time_bins, inc.nodes = FALSE,
    FADLAD = BeckLee_ages))
```
